---
title: "LP practice"
author: "Dimple"
date: "2024-09-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
```

```{r chunk_1}
data <- matrix(c(2, 30, 8, 1.5, 45, 10, 1, 40, 6), ncol = 3, byrow = TRUE)
colnames(data) <- c("material", "labour", "profit")
rownames(data) <- c("men", "women", "children")
table <- as.table(data)
table
```
  **Answer to the question**
  Suppose,\
  Number of gloves for men $$= g_{m}  $$\
  Number of gloves for women $$= g_{w}  $$\
  Number of gloves for children $$= g_{c}  $$\
  Number of full time workers $$= n_f$$\
  Number of part time workers $$= n_p$$\
  
  a. Thus the decision variables are: $$g_m, g_w, g_c, n_f, n_p$$
  b. objective function,\
  $$Max \hspace{.3cm} profit \hspace{.3cm} Z=
  8g_m+10g_w+6g_c-13(40)n_f-10(20)n_p$$
  
  c. Constraints\
  $$Material \hspace{.3cm} constraints: 2g_m+1.5g_w+g_c\le 5000$$
  
 $$Labor \hspace{.3cm} constraints: 30g_m+45g_w+40g_c\le 40n_f+20n_p$$
 
 $$full-time \hspace{.3cm}and\hspace{.3cm} part-time\hspace{.3cm}
 workers \hspace{.3cm}relationship: n_f\ge2n_p$$
 
 
 $$Non-negativity\hspace{.3cm} of\hspace{.3cm} the\hspace{.3cm}
 decision \hspace{.3cm}variables: g_m\ge0, g_w\ge0,g_c\ge0, n_f\ge0,n_p\ge0$$
 
 $$full-time \hspace{.3cm} workers \hspace{.3cm}cannot
 \hspace{.3cm}drop \hspace{.3cm}below \hspace{.3cm} 20:n_f\ge20$$
 
 
   **Solving the above LP model in R**
```{r}
library(lpSolve)
```
   
Writing down the objective function

```{r}
f.obj<-c(8,10,6,-520,-200)
```
Set constraints

```{r}
f.con<-matrix(c(2,1.5,1,0,0,
          30,45,40,-2400,-1200,
          0,0,0,-1,2),nrow = 3, byrow = TRUE)
```
Set inequality signs

```{r}
f.dir<-c("<=",
         "<=",
         "<=",
         "<=")
```
Set right hand side coefficients

```{r}
f.rhs<-c(5000,0,0,-20)
```
Get the values of obj. fun

```{r}
lp("max",f.obj,f.con,f.dir,f.rhs)
```



  
